Formalizing Stack Safety as a Security Property

### 2023-02-04 submission5/core.tex
wibbles

We propose here a formal characterization of stack safety, based on
the intuition of protecting function activations (both registers and
stack frames) from each other.
%\bcp{Aren't other formal definitions based on the same intuition?}%
%% \sna{Two differences we're trying to get across. First, protecting
%%   functions vs. protecting stack frames: we bring registers into
%%   the story because it isn't just about the stack as a data structure.
%%   Second: describing security in terms of functions' high-level
%%   security goals (f is trying to extract a secret) rather than
%%   describing operationally what they can and can't do (as the CHERI
%%   folks do.)}%
%% \bcp{Both useful!  But neither comes across, at least not very clearly.}%
We use the tools of language-based
security~\cite{sabelfeld2003language}, treating function activations as
security principals.  We decompose stack safety into a family of
properties describing
the {\em integrity} and {\em confidentiality} of the caller’s local state
and the callee's behavior during the callee's execution, plus the
{\em well-bracketed control flow} (\(\wbcf\)) property articulated by
Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
\proposecut{While our properties are motivated by the desire to specify \(LTC\) precisely,
they are stated abstractly in hopes of applying
to other enforcement mechanisms.}

@@

A {\em machine state} \(\mach \in \MACHS\) is a map from state elements to a set \(\mathcal{V}\) of
\emph{values}.
Each value \(v\) contains a \emph{payload} word, written \(|v|\).
We write \(\mach[\component]\) to denote the value of \(\mach\) at
\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
The details of value structure depend on the specific machine being modelled.
Intuitively, the payload represents the part of the value that is relevant to
the behavior of the basic machine, and it is the part that should be changed
when constructing variants (see below); the rest of the value may contain
information relevant to a hardware enforcement mechanism (such as a tag), which
should not be varied.
We assume a given \emph{compatibility} equivalence relation \(\sim\) on values,
and lift it element-wise to states.
In a tag-based enforcement scheme, values are compatible if their tags are identical.
%% \apt{This is still not very convincing!
%%   empting to say that only the payloads can affect the step function, but that is
%%   too strong, right? (OK for tags, but not, e.g. capabilities?}
%% \sna{Yeah, I don't think we can say something universal about when values should be
%%   considered compatible.}

@@

The centrality of the functional abstraction and its security are behind the
many software and hardware mechanisms proposed for its protection
\cite{Cowan+98, NagarakatteZMZ09, NagarakatteZMZ10, DeviettiBMZ08,
Kuznetsov+14, Dang+15, Shanbhogue+19, Woodruff+14, Chisnall+15,
SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises,
DBLP:conf/sp/RoesslerD18, Gollapudi+23}, which we survey in
Sections~\ref{sec:intro} and~\ref{sec:future}. Among these techniques, the micro-policies of Roessler and
DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
range of safeguards including lazy enforcement, and suitable for formulating
and validating strong security guarantees.

The only other line of work that sets out to rigorously characterize the
security of the stack is the Cerise family of program logics
\cite{SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises}.
These works define secure calling conventions for capability machines and
overlay semantics that internalize those conventions and their related stack
safety properties, phrased in terms of logical relations. Most flavors of
Cerise define experimental types of hardware capabilities in order to obtain
efficient calling conventions. Their latest paper
\cite{Georges22:TempsDesCerises} uses our work to extend their notion of stack
safety to include caller confidentiality (\(\clrc\)).

\sna{See commented out tex for content that could be massaged to fit in here,
  depending on whether we're interested in delving deeper.}\bcp{If
  there's space, some of this could be resuscitated in the related
  work section.  But removing it from here is a good call.}
%For instance, the stack-safety specification from Georges et al.
%\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
%control flow, and therefore defines {\em Well-bracketed Control Flow} (\(\wbcf\)) under the assumption that
%a callee should always return to its caller. But, in the presence of tail-call elimination, a
%return from a callee to a distant ancestor is perfectly reasonable. Similarly, the preceding work by
%Skorstengaard et al. \cite{SkorstengaardSTKJFP} made no mention of confidentiality, even though
%their enforcement mechanism did protect the confidentiality of the caller (though not the callee).

@@

HardBound \cite{DeviettiBMZ08} is another hardware extension related to
capabilities, which closely follows SoftBound \cite{NagarakatteZMZ09} and
simply aims to accelerate its enforcement of memory safety for compiled code.

@@

In \cref{sec:intro}, we discussed several dimensions of flexibility
that our properties should exhibit: applicability to different system
features and different security requirements. Another \proposecut{important dimension}
is applicability to multiple enforcement mechanisms. \apt{Could we mention that
  in intro as well?}  \proposecut{There is much future work in each of these dimensions.}

>>>>

We propose here a formal characterization of stack safety, based on
the intuition of protecting function activations (both registers and
stack frames) from each other.
%\bcp{Aren't other formal definitions based on the same intuition?}%
%% \sna{Two differences we're trying to get across. First, protecting
%%   functions vs. protecting stack frames: we bring registers into
%%   the story because it isn't just about the stack as a data structure.
%%   Second: describing security in terms of functions' high-level
%%   security goals (f is trying to extract a secret) rather than
%%   describing operationally what they can and can't do (as the CHERI
%%   folks do.)}%
%% \bcp{Both useful!  But neither comes across, at least not very clearly.}%
We use the tools of language-based
security~\cite{sabelfeld2003language}, treating function activations as
security principals.  We decompose stack safety into a family of
properties describing
the {\em integrity} and {\em confidentiality} of the caller’s local state
and the callee's behavior during the callee's execution, plus the
{\em well-bracketed control flow} (\(\wbcf\)) property articulated by
Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
\proposecut{While our properties are motivated by the desire to specify \(LTC\) precisely,
they are stated abstractly in hopes of applying
to other enforcement mechanisms.}\apt{I think we should try to keep something
  along these lines in the intro, if not this particular sentence.}

@@

A {\em machine state} \(\mach \in \MACHS\) is a map from state elements to a set \(\mathcal{V}\) of
\emph{values}.
Each value \(v\) contains a \emph{payload} word, written \(|v|\).
We write \(\mach[\component]\) to denote the value of \(\mach\) at
\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
Depending on the specific machine being modeled, values may also contain other
information relevant to hardware enforcement (such as a tag). 
%Intuitively, the payload represents the part of the value that is relevant to
%the behavior of the basic machine, and it is the part that should be changed
When constructing variants (see below), this additional information should
not be varied. To capture this idea, we assume a given \emph{compatibility} equivalence relation \(\sim\) on values,
and lift it element-wise to states.  Two values should be compatible if their
non-payload information (e.g. their tag) is identical.
%% \apt{This is still not very convincing!
%%   empting to say that only the payloads can affect the step function, but that is
%%   too strong, right? (OK for tags, but not, e.g. capabilities?}
%% \sna{Yeah, I don't think we can say something universal about when values should be
%%   considered compatible.}

@@

The centrality of the functional abstraction and its security are behind the
many software and hardware mechanisms proposed for its protection
\cite{Cowan+98, NagarakatteZMZ09, NagarakatteZMZ10, DeviettiBMZ08,
Kuznetsov+14, Dang+15, Shanbhogue+19, Woodruff+14, Chisnall+15,
SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises,
DBLP:conf/sp/RoesslerD18, Gollapudi+23}.
%, which we survey in Sections~\ref{sec:intro} and~\ref{sec:future}.
Among these techniques, the micro-policies of Roessler and
DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
range of safeguards including lazy enforcement, and suitable for formulating
and validating strong security guarantees.

To our knowledge, the only other line of work that sets out to rigorously characterize the
security of the stack is the Cerise family of program logics
\cite{SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises}.
These works define secure calling conventions for capability machines and
overlay semantics that internalize those conventions and their related stack
safety properties, phrased in terms of logical relations. Most flavors of
Cerise define experimental types of hardware capabilities in order to obtain
efficient calling conventions. Their latest paper
\cite{Georges22:TempsDesCerises} uses our work to extend their notion of stack
safety to include caller confidentiality (\(\clrc\)).

\sna{See commented out tex for content that could be massaged to fit in here,
  depending on whether we're interested in delving deeper.}\bcp{If
  there's space, some of this could be resuscitated.}
\apt{It would be very good to do this, because reviewers will want a careful comparison with this work,
  but I don't feel quite confident enough to do it (especially after my unfortunate attempt last time)!}
%For instance, the stack-safety specification from Georges et al.
%\cite{Georges22:TempsDesCerises} supposes a machine with only straightforward call-and-return
%control flow, and therefore defines {\em Well-bracketed Control Flow} (\(\wbcf\)) under the assumption that
%a callee should always return to its caller. But, in the presence of tail-call elimination, a
%return from a callee to a distant ancestor is perfectly reasonable. Similarly, the preceding work by
%Skorstengaard et al. \cite{SkorstengaardSTKJFP} made no mention of confidentiality, even though
%their enforcement mechanism did protect the confidentiality of the caller (though not the callee).

@@

\proposecut{HardBound \cite{DeviettiBMZ08} is another hardware extension related to
capabilities, which closely follows SoftBound \cite{NagarakatteZMZ09} and
simply aims to accelerate its enforcement of memory safety for compiled code.}

@@

In \cref{sec:intro}, we discussed several dimensions of flexibility
that our properties should exhibit: applicability to different system
features and different security requirements. Another 
is applicability to multiple enforcement mechanisms.
